---
title: "Choose your project"
institute: "UNSW Data Science Hub"
date: "21-11-2022"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/footer.css"]
    lib_dir: libs
    nature:
      countdown: 60000
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: true
editor_options: 
  chunk_output_type: console
---
layout: true

<div class="my-footer"><span>UNSW Data Science Hub / <a href='https://www.unsw.edu.au/research/udash'>uDASH</a></span></div>

<!-- this adds the link footer to all slides, depends on my-footer class in css-->

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "images/uDASH-logo.jpg",
  position = xaringanExtra::css_position(top = "1em", right = "1em")
)
```

```{r packages, include=FALSE}
library(fontawesome)
require(dplyr)
require(ggplot2)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_solarized_light(
#style_mono_accent(
#  base_color = "#1c5253",
  header_font_google = google_font("Roboto Slab"),
  text_font_google   = google_font("Roboto", "300", "300i"),
  code_font_google   = google_font("Roboto Mono"),
  code_font_size = '0.7rem'
)
# for ggplot2
theme_set(theme_xaringan())
```

```{r setup, include=FALSE}
# https://www.garrickadenbuie.com/blog/decouple-code-and-output-in-xaringan-slides/
knitr::opts_chunk$set(fig.dim=c(4.8, 4.5), fig.retina=2, out.width="100%")
```

---

class: inverse, center, middle


# Lifecycle of your Data Science Project

---

# Lifecycle of your Data Science Project

- Problem description & data acquisition
- Data preparation
  - Data cleaning
  - Transformation
- Exploratory data analysis
- Data modeling
- Visualisation and communication


---

class: inverse, middle


# Projects

- Psychology: Emotions during the pandemic 
- Economics: Big economic data
- Economics: Poverty
- Food and drinks: Bordeaux wines
- Ecology: Birds of New South Wales

---

class: inverse, center, middle

# Psychology: Emotions during the pandemic 

> A robust finding in emotion literature is the relationship between age and improved emotional experience... it is well-established that older individuals report more elevated emotional wellbeing than younger individuals...

***Does the emotional wellbeing advantage of older adults persisted throughout the pandemic?***

---

## Emotions ...

- Primary variables: 
  - frequency and intensity of positive emotions
  - frequency and intensity of negative emotions 
  - age
- Other variables (personality traits): 
  - Openness, 
  - Conscientiousness, 
  - Extraversion, 
  - Agreeableness and 
  - Emotional Stability


---

## Emotions ... Data

Data and metadata available here: https://osf.io/h7uqv/

You can download your file from R!

```{r}
#download.file(url="https://osf.io/download/zpfya/",
#              destfile="data/AgeAdvantagesEmotionCovid_Data.csv")

emotions_data <- read.csv("data/AgeAdvantagesEmotionCovid_Data.csv")
```


---

## Emotions ... Cleaning and transformation

Select and rename your variables

```{r}

tibble(emotions_data)
```

---

## Emotions ... Explore

Range of ages, mean and extreme values of intensity and frequency of emotions, personality traits 

.pull-left[
```{r emo-eda, eval=FALSE}
ggplot(data=emotions_data, 
       aes(x=age)) + 
  geom_histogram(bins=14)
```
]

.pull-right[
```{r emo-eda-out, ref.label="emo-eda", echo=FALSE, warning=FALSE, message=FALSE}
```
]


---

## Emotions ... Model

Increase or decrease of emotions by age? differences between groups?

```{r output-}
model <- lm(avg_i_neg ~ age,
   emotions_data)
summary(model)
```

---

## Emotions ... Visualise

Combine scatterplot and regression lines

.pull-left[
```{r emo-viz, eval=FALSE}
ggplot(data=emotions_data, 
       aes(x=age,y=avg_i_neg)) + 
  geom_point() + 
  geom_smooth(method = lm)
```
]

.pull-right[
```{r emo-viz-out, ref.label="emo-viz", echo=FALSE, warning=FALSE,message=FALSE}
```
]

---

# Thanks!

[`r fa(name = "twitter")` @uDASH_UNSW](https://twitter.com/uDASH_UNSW)

[`r fa(name = "link")` //www.unsw.edu.au/research/udash](https://www.unsw.edu.au/research/udash)

[`r fa(name = "paper-plane")` uDASH@unsw.edu.au](mailto:uDASH@unsw.edu.au )

