---
title: "Choose your project"
institute: "UNSW Data Science Hub"
date: "21-11-2022"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/footer.css"]
    lib_dir: libs
    nature:
      countdown: 60000
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: true
editor_options: 
  chunk_output_type: console
---
layout: true

<div class="my-footer"><span>UNSW Data Science Hub / <a href='https://www.unsw.edu.au/research/udash'>uDASH</a></span></div>

<!-- this adds the link footer to all slides, depends on my-footer class in css-->

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "images/uDASH-logo.jpg",
  position = xaringanExtra::css_position(top = "1em", right = "1em")
)
```

```{r packages, include=FALSE}
library(fontawesome)
require(dplyr)
require(ggplot2)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_solarized_light(
#style_mono_accent(
#  base_color = "#1c5253",
  header_font_google = google_font("Roboto Slab"),
  text_font_google   = google_font("Roboto", "300", "300i"),
  code_font_google   = google_font("Roboto Mono"),
  code_font_size = '0.7rem'
)
# for ggplot2
theme_set(theme_xaringan())
```

```{r setup, include=FALSE}
# https://www.garrickadenbuie.com/blog/decouple-code-and-output-in-xaringan-slides/
knitr::opts_chunk$set(fig.dim=c(4.8, 4.5), fig.retina=2, out.width="100%")
```


---

class: inverse, center, middle

# Wineinformatics: Bordeaux wines

> Bordeaux, France, is considered as the most famous wine region in the world... _Wineinformatics_ is the quantitative analysis of wine reviews... 

***What is the relationship between wine critic ratings and wine prices?***

---

## Bordeaux wines ...

We look at Bordeaux wine prices in the UK, and the ranking of two critics (Robert Parker from the US and Clive Coates from the UK).

- Variables of interest: 
  - price, 
  - ratings by each critic
- Other variables: 
  - other indicators of quality

---

## Bordeaux wines ... Data

Data provided by uDASH staff (but other [datasets available online](https://ieee-dataport.org/open-access/wineinformatics-21st-century-bordeaux-wines-dataset))


```{r}
wine_data <- read.csv("data/Bordeaux.csv", header=TRUE)
head(wine_data)

```


---

## Bordeaux wines ... Cleaning and transformation

Logarithmic transformation

```{r}
wine_data$logPrice <- log(wine_data$Price)
summary(wine_data[,c("Price","logPrice")])
```

---

## Bordeaux wines ... Explore

Scatterplots and boxplots of variables

.pull-left[
```{r emo-eda, eval=FALSE}
ggplot(wine_data) +
  geom_point(aes(ParkerPoints, 
                 CoatesPoints,
                 color=logPrice)) +
  xlab("Parker's rating") +
  ylab("Coate's rating") +
  scale_xaringan_color_continuous()
```
]

.pull-right[
```{r emo-eda-out, ref.label="emo-eda", echo=FALSE, warning=FALSE, message=FALSE}
```
]


---

## Bordeaux wines ... Model

Linear regression, logistic regression

```{r output-}
model <- glm(Price ~ ParkerPoints + Pomerol, data=wine_data)
summary(model)
```

---

## Bordeaux wines ... Visualise

Combine scatterplot and regression lines

.pull-left[
```{r wine-viz, eval=FALSE}
ggplot(wine_data,
       aes(ParkerPoints, logPrice)) +
  geom_point() +
  xlab("Parker's rating") +
  ylab("log(Price) (log-pounds)") +
  geom_smooth(method="lm")
```
]

.pull-right[
```{r wine-viz-out, ref.label="wine-viz", echo=FALSE, warning=FALSE,message=FALSE}
```
]

---


# Thanks!

[`r fa(name = "twitter")` @uDASH_UNSW](https://twitter.com/uDASH_UNSW)

[`r fa(name = "link")` //www.unsw.edu.au/research/udash](https://www.unsw.edu.au/research/udash)

[`r fa(name = "paper-plane")` uDASH@unsw.edu.au](mailto:uDASH@unsw.edu.au )

